Quality Criteria for SKOS Vocabularies

### 2012-03-21 criteria.tex
resolved some todos

\begin{definition}[SKOS Vocabulary] Let a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV \rangle$, with \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) being the set of \textbf{resources}, \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) being the set of \textbf{conceptual resources}, $LV = I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$ being the set of \todo{CM}{\textbf{plain literals}?}, and \item \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) being the set of \textbf{semantic relations} associating conceptual resources with one another. Further, we let $V$ be the fully entailed RDFS interpretation~\cite{RDFSEM2012} of the underlying RDF graph, enriched by entailment of \texttt{owl:inverseOf} properties.

@@

\todo{All}{decide for either using "concept" or "conceptual resource"}

@@

This criterion is an extension of a statement from \cite{Isaac2009}, recommending ``no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme''. Based on our review of existing vocabularies, we could observe that (i) many vocabularies don’t make use of concept schemes and (ii) also identical \texttt{skos:altLabels} or \texttt{skos:hiddenLabels} can have a negative effect in some use-cases (e.g, auto-completion of concept labels based on user input). Thus we follow the more general approach of finding all pairs of concepts with their respective \texttt{skos:prefLabel}, \texttt{skos:altLabel} or \texttt{skos:hiddenLabel} meeting a threshold of a certain similarity function $sim:LV \times LV \rightarrow [0,1]$. For the analysis of the criterion in this document, we define the similarity function to check for case-insensitive string equality and the threshold to equal 1. Calculation of this criterion is done by iterating over all pairs of concepts, i.e., elements of $C \times C$. We then apply the $sim$ to every possible combination of labels between the members of the pairs and collect those pairs for which $sim$ meets or exceeds a specified threshold.

@@

\subsubsection{Omitted Top Concepts}
The SKOS language provides ConceptSchemes which are a facility for grouping related concepts. In order to provide entry points to such a group of concepts, one or more concepts can be marked as top concepts. This helps to provide ``efficient access'' (SKOS primer) and simplifies orientation in the vocabulary. To calculate this criterion, we identify all ConceptSchemes $I_{CEXT}(\texttt{skos:ConceptScheme}^\mathcal{I})$ not contained in any of the pairs in $I_{EXT}(\texttt{skos:hasTopConcept}^\mathcal{I})$.

>>>>

\begin{definition}[SKOS Vocabulary] Let a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, ULV \rangle$, with \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) being the set of \textbf{resources}, \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) being the set of \textbf{conceptual resources}, $ULV \subseteq I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$ being the set of \textbf{untyped literal values}, and \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) being the set of \textbf{semantic relations} associating conceptual resources with one another. Further, we let $V$ be the fully entailed RDFS interpretation~\cite{RDFSEM2012} of the underlying RDF graph, enriched by entailment of \texttt{owl:inverseOf} properties.

@@

\todo{All}{decide for either using the term "concept" or "conceptual resource"}

@@

This criterion is an extension of a statement from \cite{Isaac2009}, recommending ``no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme''. Based on our review of existing vocabularies, we could observe that (i) many vocabularies don’t make use of concept schemes and (ii) also identical \texttt{skos:altLabels} or \texttt{skos:hiddenLabels} can have a negative effect in some use-cases (e.g, auto-completion of concept labels based on user input). Thus we follow the more general approach of finding all pairs of concepts with their respective \texttt{skos:prefLabel}, \texttt{skos:altLabel} or \texttt{skos:hiddenLabel} meeting a threshold of a certain similarity function $sim:ULV \times ULV \rightarrow [0,1]$. For the analysis of the criterion in this document, we define the similarity function to check for case-insensitive string equality and the threshold to equal 1. Calculation of this criterion is done by iterating over all pairs of concepts, i.e., elements of $C \times C$. We then apply the $sim$ to every possible combination of labels between the members of the pairs and collect those pairs for which $sim$ meets or exceeds a specified threshold.

@@

\subsubsection{Omitted Top Concepts}
The SKOS language provides ConceptSchemes which are a facility for grouping related concepts. In order to provide entry points to such a group of concepts, one or more concepts can be marked as top concepts. This helps to provide ``efficient access''~\cite{Isaac2009} and simplifies orientation in the vocabulary. To calculate this criterion, we identify all ConceptSchemes $I_{CEXT}(\texttt{skos:ConceptScheme}^\mathcal{I})$ not contained in any of the pairs in $I_{EXT}(\texttt{skos:hasTopConcept}^\mathcal{I})$.
